Bayesian non-parametric inference for multivariate peaks-over-threshold models

### 2023-05-08 introduction.tex
before some cutting

To develop a multivariate PoT model for extreme values consider a 
$d$-dimensional random vector $\bm{W} = (W_1, \ldots ,W_d)$ with
cumulative distribution $F$. Following \cite{RoSeWa2018a}, assume 
that there exists sequences of vectors $\bm{a}_n$ and $\bm{b}_n$,
and a $d$-variate distribution $G$ such that 
$\lim_{n\rightarrow\infty} F^n(\bm{a}_n \bm{w} + \bm{b}_n) = 
G(\bm{w})$. $G$ is a $d$-variate generalized extreme value 
distribution. It follows that
\begin{equation*}
    \begin{aligned}
    \lim\limits_{n\rightarrow\infty} &\prob{\bm{a}^{-1}_n (\bm{W} - \bm{b}_n) 
   \leq \bm{w} \mid \bm{W} \not\leq \bm{b}_n)} \\ 
     &\hspace{0.5cm}= \frac{\log G(\bm{w} \wedge \bm{0}) 
    - \log G(\bm{w}) }{\log G(\bm{0})} = H(\bm{w}).
    \end{aligned}
\end{equation*}
$H$ is the multivariate Pareto distribution. \cite{RoSeWa2018a}
provide a number of stochastic representations for $H$. In particular
Remark 1 justifies the representation given in \cite{ferreira2014},
consisting of taking $\bm{W}$, in the limit, as $\bm{W} = \bm{V}R$ 
where $R$ and $\bm{V}$ are independent. $R = ||\bm{W}||_\infty$ is 
distributed as a standard Pareto  random variable, and $\bm{V} = 
\bm{W}/||\bm{W}||_\infty$ is a random vector in  
$\mathbb{S}_{\infty}^{d-1}$, the positive orthant of the 
unit  sphere in infinite norm, with distribution $\Phi$. 
%More  specifically, 
%using the notation in Theorem 6 of \cite{RoSeWa2018a}, define $\bm{S} 
%=  \log(\bm{V})$.  Then, according to Definition 5, $\bm{S}$ 
%is a spectral  random vector. Taking  $E \sim \text{Exp}(1)$, 
%independent of  $\bm{V}$,  we have that $\bm{Y} = \bm{S} + E$  
%follows a multivariate GP, according  to Theorem 6.  Then, using 
%Proposition 4,  $\bm{V}R = e^{\bm{Y}}$ will also follow
%a multivariate GP. Notice that $R = e^E$ is distributed as 
%a standard Pareto random variable. 
$R$ and $\bm{V}$ are referred, respectively, as the {\em radial} 
and {\em angular}  components of $H$. The angular measure controls 
the dependence  structure of $\bm{W}$  in the tails. In view of 
this, to obtain a  PoT model we seek a flexible model for the 
distribution of $\bm{V} \in {\mathbb S}_{\infty}^{d-1}$.

%\subection{Pairwise Extremal Dependence\label{subsec:ped}}
A measure that is used to characterize the strength of the dependence,
in the tail, for two random variables $Z_1$ and $Z_2$, with marginal
distributions $F_1$ and $F_2$ is given by \citep{coles2001}.
\[
    \chi_{12} = \lim\limits_{u\uparrow 1} \prob{F_1(Z_1)>u\mid F_2(Z_2)>u}  
\]
$\chi_{12}$ provides information about the distribution of extremes for the variable $Z_1$
given that $Z_2$ is very large.  When $\chi_{12}>0$, $Z_1$ and $Z_2$ 
are said to
be asymptotically dependent, otherwise they are asymptotically
independent. The following result provides the asymptotic dependence coefficient between two components of $\bm{Z}$
after PoT limit. 
\begin{prop}\label{ppchi}
Suppose that $\bm{Z} = R\bm{V}$ with $R\sim Pa(1)$,
$\prob{V_\ell > 0} = 1$ and $\expect{V_\ell}$ exists, for $\ell=1, \ldots ,d$, then
\begin{equation}
    \label{eqn:chi_ij}
	\chi_{\jmath\ell} = \expect{\frac{V_\jmath}{\expect{V_\jmath}} \wedge \frac{V_\ell}{\expect{V_\ell}}}
\end{equation}
\end{prop}
{\em Proof:}
Denote as $F_\ell$ the marginal distribution of $Z_\ell$. To obtain $\chi_{\jmath\ell}$ we need $Pr(Z_\jmath>z_\jmath,Z_\ell>z_\ell)$, where $z_\ell =
F_\ell^{-1}(u) = \expect{V_\ell}/(1 - u), \;\ell=1,\dots, d$.
Using the fact that
$V_\ell>0, \forall \ell$ almost surely, we have that the former is equal to
\begin{equation*}
    \begin{aligned}
    &\hspace{-0.5cm}\prob{R>\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}}
    = \expect{1\wedge\left(\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}\right)^{-1}}\\
    &= \expect{\frac{V_\jmath}{z_\jmath}\wedge\frac{V_\ell}{z_\ell}}
    = (1 - u)\expect{\frac{V_\jmath}{\expect{V_\jmath}}\wedge\frac{V_\ell}{\expect{V_\ell}}},
    \end{aligned}
\end{equation*}
where the second identity is justified by the fact that $V_i$ is
bounded and $z_i\rightarrow\infty$. The proof is completed by noting
that $\prob{F_i(Z_i)>u} = 1 - u. \hfill\Box$

%\subsection{Multivariate Conditional Survival Function\label{subsec:condsurv}}
Equation \ref{eqn:chi_ij} implies that $\chi_{\jmath\ell}>0$, and so, no asymptotic independence is possible under our proposed model. For the analysis of extreme values it is of interest to calculate the multivariate conditional survival function. The following result provides the relevant expression, as a function of the angular measure.
\begin{prop}
Assume the same conditions of Proposition \ref{ppchi}. 
Let $\alpha \subset \{1, \ldots ,d\}$ be a collections of indexes. 
Then     
\begin{equation} \label{eqn:condsurv}
    \begin{aligned}
    &\hspace{-0.5cm}\mathrm{Pr}\left[\bigcap_{\ell \in \alpha} Z_\ell > z_\ell \mid \bigcap_{\ell\not\in\alpha} Z_\ell > z_\ell\right] \\
    &= \frac{\expect{\bigwedge_{k = 1}^d 1\wedge \frac{V_k}{z_k}}}{\expect{\bigwedge_{k \not\in\alpha}1\wedge\frac{V_k}{z_k}}}.
    \end{aligned}
  \end{equation}
\end{prop}  
The proof uses a similar approach to the proof of Proposition \ref{ppchi}.

  % EOF

>>>>

To develop a multivariate PoT model for extreme values consider a 
$d$-dimensional random vector $\bm{W} = (W_1, \ldots ,W_d)$ with
cumulative distribution $F$. Following \cite{RoSeWa2018a}, assume 
that there exists sequences of vectors $\bm{a}_n$ and $\bm{b}_n$,
and a $d$-variate distribution $G$ such that 
$\lim_{n\rightarrow\infty} F^n(\bm{a}_n \bm{w} + \bm{b}_n) = 
G(\bm{w})$. $G$ is a $d$-variate generalized extreme value 
distribution. It follows that
\begin{equation*}
    \begin{aligned}
    \lim\limits_{n\rightarrow\infty} &\prob{\bm{a}^{-1}_n (\bm{W} - \bm{b}_n) 
   \leq \bm{w} \mid \bm{W} \not\leq \bm{b}_n)} \\ 
     &\hspace{0.5cm}= \frac{\log G(\bm{w} \wedge \bm{0}) 
    - \log G(\bm{w}) }{\log G(\bm{0})} = H(\bm{w}).
    \end{aligned}
\end{equation*}
$H$ a multivariate Pareto distribution, and corresponds to the joint
distribution conditional on exceeding a multivariate threshold. In addition to
a copula function, $H$ depends on two $d$-dimensional vectors of 
parameters that we denote as $\bm{\xi}$ for the shapes and $\bm{\sigma}$ 
for the scales. \cite{RoSeWa2018a} provide a number of stochastic 
representations for $H$.  In particular, denote as $Z$ a
random variable with distribution $H$ where $\bm{\xi}= 1$ and 
$\bm{\sigma} = 0$.  Then, Remark 1 justifies the representation in 
\cite{ferreira2014}, giving $\bm{Z} = \bm{V}R$
where $R$ and $\bm{V}$ are independent. $R = \|\bm{Z}\|_\infty$ is 
distributed as a standard Pareto random variable, and $\bm{V} = 
\bm{Z}/\|\bm{Z}\|_\infty$ is a random vector in  
$\mathbb{S}_{\infty}^{d-1}$, the positive orthant of the 
unit sphere under $\mathcal{L}_{\infty}$ norm, with distribution $\Phi$. 
This representation is central to the methods proposed in this paper.
$R$ and $\bm{V}$ are referred to, respectively, as the {\em radial} 
and {\em angular} components of $H$. The angular measure controls 
the dependence  structure of $\bm{Z}$  in the tails. In view of 
this, to obtain a  PoT model we seek a flexible model for the 
distribution of $\bm{V} \in {\mathbb S}_{\infty}^{d-1}$.

The approach considered in \cite{rootzen2018} focuses on the limiting
conditional distribution $H$. An alternative approach consists of assuming
that regular variation \cite[see, for example,][]{resnick2008extreme} holds for 
the limiting distribution of $\bm{W}$, implying
\[
    \lim\limits_{n\rightarrow\infty} n \prob{n^{-1} \bm{W}\in A} = 
    \mu(A),
\]
where $\mu$ is termed the exponent measure. $\mu$ has the homogeneity property
$\mu(tA) = t^{-1}\mu(A)$. Letting $\rho = \|W\|_p, p>0$ 
and $\bm{\theta} = \bm{W}/\rho$, define  $\Psi(B) = \mu(\{\bm{w} : \rho>1, 
\bm{\theta} \in B\})$, which is referred to as the angular measure. After 
some manipulations, we obtain that
\begin{equation}
\label{exponent-mu}
    \lim\limits_{r\rightarrow\infty} \text{Pr}\left[\bm{\theta}\in A \rvert \rho>r\right] = 
    \frac{\Psi(A)}{\Psi({\mathbb S}_p^{d-1})} .
\end{equation}
Thus, a model for the exponent measure induces a model for the limiting distribution
conditional on the observations being above a threshold defined with respect
to their $p$-norm. The constraint that all marginals of $\mu$ correspond to a
standard Pareto distribution leads to the so called moment constraints on $\Psi$. 
Inference for the limiting distribution of the exceedances
needs to account for the normalizing constant in Equation \ref{exponent-mu}, as well
the moment constraints. In addition, the moment constraints can impose unrealistic
symmetries in the data distribution.

%\subection{Pairwise Extremal Dependence\label{subsec:ped}}
%%A measure that is used to characterize the strength of the dependence,
%%in the tail, for two random variables $Z_1$ and $Z_2$, with marginal
%%distributions $F_1$ and $F_2$ is given by \citep{coles2001}.
%%\[
%%    \chi_{12} = \lim\limits_{u\uparrow 1} \prob{F_1(Z_1)>u\mid F_2(Z_2)>u}  
%%\]
%%$\chi_{12}$ provides information about the distribution of extremes for the variable $Z_1$
%%given that $Z_2$ is very large.  When $\chi_{12}>0$, $Z_1$ and $Z_2$ 
%%are said to
%%be asymptotically dependent, otherwise they are asymptotically
%%independent. The following result provides the asymptotic dependence coefficient between two components of $\bm{Z}$
%%after PoT limit. 
%%\begin{prop}\label{ppchi}
%%Suppose that $\bm{Z} = R\bm{V}$ with $R\sim Pa(1)$,
%%$\prob{V_\ell > 0} = 1$ and $\expect{V_\ell}$ exists, for $\ell=1, \ldots ,d$, then
%%\begin{equation}
%%    \label{eqn:chi_ij}
%%	\chi_{\jmath\ell} = \expect{\frac{V_\jmath}{\expect{V_\jmath}} \wedge \frac{V_\ell}{\expect{V_\ell}}}
%%\end{equation}
%%\end{prop}
%%{\em Proof:}
%%Denote as $F_\ell$ the marginal distribution of $Z_\ell$. To obtain $\chi_{\jmath\ell}$ we need $Pr(Z_\jmath>z_\jmath,Z_\ell>z_\ell)$, where $z_\ell =
%%F_\ell^{-1}(u) = \expect{V_\ell}/(1 - u), \;\ell=1,\dots, d$.
%%Using the fact that
%%$V_\ell>0, \forall \ell$ almost surely, we have that the former is equal to
%%\begin{equation*}
%%    \begin{aligned}
%%    &\hspace{-0.5cm}\prob{R>\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}}
%%    = \expect{1\wedge\left(\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}\right)^{-1}}\\
%%    &= \expect{\frac{V_\jmath}{z_\jmath}\wedge\frac{V_\ell}{z_\ell}}
%%    = (1 - u)\expect{\frac{V_\jmath}{\expect{V_\jmath}}\wedge\frac{V_\ell}{\expect{V_\ell}}},
%%    \end{aligned}
%%\end{equation*}
%%where the second identity is justified by the fact that $V_i$ is
%%bounded and $z_i\rightarrow\infty$. The proof is completed by noting
%%that $\prob{F_i(Z_i)>u} = 1 - u. \hfill\Box$
%%
%%%\subsection{Multivariate Conditional Survival Function\label{subsec:condsurv}}
%%Equation \ref{eqn:chi_ij} implies that $\chi_{\jmath\ell}>0$, and so, no asymptotic independence is possible under our proposed model. For the analysis of extreme values it is of interest to calculate the multivariate conditional survival function. The following result provides the relevant expression, as a function of the angular measure.
%%\begin{prop}
%%Assume the same conditions of Proposition \ref{ppchi}. 
%%Let $\alpha \subset \{1, \ldots ,d\}$ be a collections of indexes. 
%%Then     
%%\begin{equation} \label{eqn:condsurv}
%%    \begin{aligned}
%%    &\hspace{-0.5cm}\mathrm{Pr}\left[\bigcap_{\ell \in \alpha} Z_\ell > z_\ell \mid \bigcap_{\ell\not\in\alpha} Z_\ell > z_\ell\right] \\
%%    &= \frac{\expect{\bigwedge_{k = 1}^d 1\wedge \frac{V_k}{z_k}}}{\expect{\bigwedge_{k \not\in\alpha}1\wedge\frac{V_k}{z_k}}}.
%%    \end{aligned}
%%  \end{equation}
%%\end{prop}  
%%The proof uses a similar approach to the proof of Proposition \ref{ppchi}.
%%
  % EOF

### 2023-05-08 methodology_projgamma.tex
before some cutting

\subsection{Projected gamma family\label{subsec:projgamma}}
  A natural  distribution to consider in ${\mathbb R}^d_+$ is given by a product of independent
  univariate Gamma distributions. Let
    $\bm{ X} \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{\ell}\mid\alpha_{\ell},\beta_{\ell}\right)$, 
    $\alpha_\ell$ and $\beta_\ell$ are the shape and scale parameters, respectively. Using Equations (\ref{eqn:pnormt}) -- (\ref{eqn:pnormjac}) we have the joint density
  \begin{equation}
  \label{joint}
    \begin{aligned}
    f(r,\bm{ y}) = &\prod_{\ell = 1}^{d}
      \left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}(ry_{\ell})^{\alpha_{\ell} - 1}
          \exp\lbrace-\beta_{\ell}ry_{\ell}\rbrace\right] \\
      &\hspace{0.5cm}\times r^{d-1}\left[y_d +
            {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right].
    \end{aligned}
  \end{equation}
  Integrating out $r$ yields the resulting \emph{Projected Gamma} density
  \begin{equation}
    \label{eqn:projgamma}
    \begin{aligned}
    \text{PG}(\bm{ y}\mid\bm{ \alpha},\bm{ \beta}) &=
          \prod_{\ell = 1}^d\left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}
                y_{\ell}^{\alpha_{\ell} - 1}\right]\\
      &\hspace{0.5cm}\times \left[y_d +
          {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right]\\
      &\hspace{0.5cm}\times \frac{\Gamma({\textstyle\sum}_{\ell = 1}^d\alpha_{\ell})}{\left({\textstyle\sum}_{\ell = 1}^d
                    \beta_{\ell}y_{\ell}\right)^{{\scriptstyle\sum_{\ell = 1}^d \alpha_{\ell}}}},
    \end{aligned}
  \end{equation}
  defined for $\bm{y}\in {\mathbb S}_p^{d-1}$, and for any $p>0$. To avoid identifiability problems when estimating the shape and scale parameters, we set $\beta_1 = 1$.
  \cite{nunez2019} obtain the density in Equation (\ref{eqn:projgamma})
  for $p=2$ as a multivariate distribution for directional data, using spherical coordinates.  
  For $\bm{ y}\in {\mathbb S}_1^{d-1}$, and $\beta_{\ell} = \beta$ for all $\ell$, the density in Equation (\ref{eqn:projgamma}) corresponds to that of a Dirichlet distribution.

@@

>>>>

\subsection{Projected gamma family\label{subsec:projgamma}}
To obtain a distribution on 
${\mathbb S}_{p}^{d-1}=\{\bm{s} : \bm{s} \in {\mathbb R}_{+}^{d}, \| \bm{s}\|_{p} = 1\}$ 
we start with a vector in 
$\bm{x} \in {\mathbb R}^d_+$, calculate its $\mathcal{L}_p$-norm  defined as
  \begin{equation*}
    \lVert \bm{x} \rVert_p = 
    \left({\textstyle\sum}_{\ell = 1}^d \lvert s_{\ell}\rvert^p\right)^{\frac{1}{p}},
  \end{equation*}
  and normalize it to obtain $\bm{y} = \bm{x}/\|\bm{x}\|_p \in {\mathbb S}_p^{d-1}$. 
  The absolute and Euclidean norms are obtained for $p=1$ and $p=2$ respectively,
  and the $\mathcal{L}_{\infty}$ norm can be obtained as a limit, 
  \begin{equation*}
    \| \bm{s} \|_{\infty}
      = \lim\limits_{p\to\infty} \| \bm{s} \|
      = \underset{\ell = 1}{\overset{d}{\bigvee}}s_{\ell}.
  \end{equation*}

  A natural  distribution to consider in ${\mathbb R}^d_+$ is given by a product of independent
  univariate Gamma distributions. Let
    $\bm{ X} \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{\ell}\mid\alpha_{\ell},\beta_{\ell}\right)$, 
    $\alpha_\ell$ and $\beta_\ell$ are the shape and scale parameters, respectively. 
    For any finite $p$, letting 
   $ y_d = (1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p)^{1/p}$,
    the transformation
  \begin{equation}
    \label{eqn:pnormt}
    \begin{aligned}
    T(x_1,\ldots,x_d) &= \left(\pnorm{\bm{x}}{p}, \frac{x_1}{\pnorm{\bm{x}}{p}},
                          \ldots , \frac{x_{d-1}}{\pnorm{\bm{x}}{p}}\right) \\
                        &= (r,y_1,\ldots,y_{d-1})
    \end{aligned}
  \end{equation}
  is invertible with
  \begin{equation}
    \label{eqn:pnormtinv}
    \begin{aligned}
    &\hspace{-0.5cm}T^{-1}\left(r,y_1,\ldots,y_{d-1}\right)\\
      & = \left(ry_1,\ldots,ry_{d-1}, r\left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}}\right).
    \end{aligned}
  \end{equation}
  The Jacobian of the transformation takes the form
  \begin{equation}
    \label{eqn:pnormjac}
    \begin{aligned}
    r^{d-1}\left[\left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}} \right. & \\
        &\hspace{-2cm}+ \left. {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(1 - {\textstyle\sum}_{l=1}^{d-1} y_{\ell}^p\right)^{\frac{1}{p} - 1}\right].
    \end{aligned}
  \end{equation}
  The normalization provided by $T$ maps a vector in ${\mathbb R}_+^d$ onto
  ${\mathbb S}_p^{d-1} $. With a slight abuse of terminology we refer to it as a
  projection. Using Equations (\ref{eqn:pnormt}) -- (\ref{eqn:pnormjac}) we have the joint density
  \begin{equation}
  \label{joint}
    \begin{aligned}
    f(r,\bm{ y}) = &\prod_{\ell = 1}^{d}
      \left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}(ry_{\ell})^{\alpha_{\ell} - 1}
          \exp\lbrace-\beta_{\ell}ry_{\ell}\rbrace\right] \\
      &\hspace{0.5cm}\times r^{d-1}\left[y_d +
            {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right].
    \end{aligned}
  \end{equation}
  Integrating out $r$ yields the resulting \emph{Projected Gamma} density
  \begin{equation}
    \label{eqn:projgamma}
    \begin{aligned}
    \text{PG}(\bm{ y}\mid\bm{ \alpha},\bm{ \beta}) &=
          \prod_{\ell = 1}^d\left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}
                y_{\ell}^{\alpha_{\ell} - 1}\right]\\
      &\hspace{0.5cm}\times \left[y_d +
          {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right]\\
      &\hspace{0.5cm}\times \frac{\Gamma({\textstyle\sum}_{\ell = 1}^d\alpha_{\ell})}{\left({\textstyle\sum}_{\ell = 1}^d
                    \beta_{\ell}y_{\ell}\right)^{{\scriptstyle\sum_{\ell = 1}^d \alpha_{\ell}}}},
    \end{aligned}
  \end{equation}
  defined for $\bm{y}\in {\mathbb S}_p^{d-1}$, and for any finite $p>0$. To avoid identifiability 
  problems when estimating the shape and scale parameters, we set $\beta_1 = 1$.
  \cite{nunez2019} obtain the density in Equation (\ref{eqn:projgamma})
  for $p=2$ as a multivariate distribution for directional data, using spherical coordinates.  
  For $\bm{ y}\in {\mathbb S}_1^{d-1}$, and $\beta_{\ell} = \beta$ for all $\ell$, the density in Equation (\ref{eqn:projgamma}) corresponds to that of a Dirichlet distribution.

@@

A measure that is used to characterize the strength of the dependence,
  in the tail, for two random variables $Z_1$ and $Z_2$, with marginal
  distributions $F_1$ and $F_2$ is given by \citep{coles2001}
  \[
    \chi_{12} = \lim\limits_{u\uparrow 1} \prob{F_1(Z_1)>u\mid F_2(Z_2)>u}.   
  \]
  $\chi_{12}$ provides information about the distribution of extremes for the variable $Z_1$
  given that $Z_2$ is very large.  When $\chi_{12}>0$, $Z_1$ and $Z_2$ are said to
  be asymptotically dependent, otherwise they are asymptotically independent. The following result 
  provides the asymptotic dependence coefficient between two components of $\bm{Z}$ after PoT limit.
  \begin{prop}\label{ppchi}
  Suppose that $\bm{Z} = R\bm{V}$ with $R\sim Pa(1)$,
  $\prob{V_\ell > 0} = 1$ and $\expect{V_\ell}$ exists, for $\ell=1, \ldots ,d$, then
    \begin{equation}
    \label{eqn:chi_ij}
	\chi_{\jmath\ell} = \expect{\frac{V_\jmath}{\expect{V_\jmath}} \wedge \frac{V_\ell}{\expect{V_\ell}}}
    \end{equation}
  \end{prop}
  {\em Proof:}
  Denote as $F_\ell$ the marginal distribution of $Z_\ell$. To obtain $\chi_{\jmath\ell}$ we need 
  $Pr(Z_\jmath>z_\jmath,Z_\ell>z_\ell)$, where $z_\ell = F_\ell^{-1}(u) = \expect{V_\ell}/(1 - u), \;\ell=1,\dots, d$.  
  Using the fact that $V_\ell>0, \forall \ell$ almost surely, we have that the former is equal to
  \begin{equation*}
    \begin{aligned}
    &\hspace{-0.5cm}\prob{R>\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}}
    = \expect{1\wedge\left(\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}\right)^{-1}}\\
    &= \expect{\frac{V_\jmath}{z_\jmath}\wedge\frac{V_\ell}{z_\ell}}
    = (1 - u)\expect{\frac{V_\jmath}{\expect{V_\jmath}}\wedge\frac{V_\ell}{\expect{V_\ell}}},
    \end{aligned}
  \end{equation*}
  where the second identity is justified by the fact that $V_i$ is bounded and $z_i\rightarrow\infty$. The 
  proof is completed by noting that $\prob{F_i(Z_i)>u} = 1 - u. \hfill\Box$

Equation \ref{eqn:chi_ij} implies that $\chi_{\jmath\ell}>0$, and so, no asymptotic independence is possible 
  under our proposed model. For the analysis of extreme values it is of interest to calculate the multivariate 
  conditional survival function. The following result provides the relevant expression, as a function of the 
  angular measure.
  \begin{prop}
  Assume the same conditions of Proposition \ref{ppchi}.   Let $\alpha \subset \{1, \ldots ,d\}$ be a 
  collections of indexes.  Then     
  \begin{equation} \label{eqn:condsurv}
    \begin{aligned}
    &\hspace{-0.5cm}\mathrm{Pr}\left[\bigcap_{\ell \in \alpha} 
        Z_\ell > z_\ell \mid \bigcap_{\ell\not\in\alpha} Z_\ell > z_\ell\right] \\
    &= \frac{\expect{\bigwedge_{k = 1}^d 1\wedge \frac{V_k}{z_k}}}{\expect{\bigwedge_{k \not\in\alpha}1\wedge\frac{V_k}{z_k}}}.
    \end{aligned}
  \end{equation}
  \end{prop}  
  The proof uses a similar approach to the proof of Proposition \ref{ppchi}.

Equations \ref{eqn:chi_ij} and \ref{eqn:condsurv} provide relevant tools for inference on the tail 
  behavior of the joint distribution of the observations. The expressions can be readily calculated 
  within a sample-based inferential approach like the one considered in the following section.
